---
title: "Increasing clutch size 2and success"
subtitle: "draft results" 
author:     | 
    | Quinlan Mann
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document:
    toc: no
  word_document:
    toc: no
  html_document:
    toc: no
    keep_md: yes
md_document:
    variant: markdown_github
header-includes:
  - \usepackage{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \AtBeginEnvironment{tabular}{\singlespacing}
  - \AtBeginEnvironment{lltable}{\singlespacing}
  - \AtBeginEnvironment{tablenotes}{\doublespacing}
linestretch: 1
always_allow_html: true
---

```{r, include=F, echo=F, warning=F}
library(tidyverse)
library(glmmTMB)
library(gridExtra)
library(sjPlot)
library(broom.mixed)
library(bbmle)
library(performance)
library(see)
library(ggthemes)
library(reshape2)
library(ggeffects)
```

```{r, setup, include=F}
knitr::opts_chunk$set(echo = F, fig.width=6, fig.height=4, warning = F, message = F)
options(digits = 3)
set.seed(19970227)
```

```{r, include=F}
#data import and cleanup, ensuring that all columns are appropriate types
ANI<-(read.csv("E:/quinl/Desktop/Grad School/ANI/NEST_SUMMARY_ANIS.csv")	%>% 
        mutate(NULL, 
               ADULTS=as.numeric(ADULTS), 
               EGGS_BURIED=as.numeric(EGGS_BURIED),
               EGGS_BURIED=as.numeric(EGGS_BURIED), 
               EGGS_UNBURIED=as.numeric(EGGS_UNBURIED),
               HATCHED=as.numeric(HATCHED), 
               FLEDGED=as.numeric(FLEDGED),
               TOT_EGGS=as.numeric(TOT_EGGS),
               LOCATION=as.factor(LOCATION),
               YEAR=as.factor(YEAR), 
               SITE=as.factor(SITE)	
        ))

#removes the missing values
ANI<- ANI[!(is.na(ANI$ADULTS)), ] 
ANI<- ANI[! (ANI$ADULTS==1), ] #removes the record of 1 adult from the dataset
ANI<-ANI %>% #removes the unknown groups
  group_by(SITE) %>%
  filter(n() > 12)

#removing Depredated nests from dataset and notes columns
ANI<-subset(ANI, NEST.FATE!="DP")[,c(1:12)]
ANI$COMP<-ANI$EGGS_UNBURIED/ANI$TOT_EGGS

#setting the theme for ggplots
theme_set(ggthemes::theme_few())
```

```{r}

success_mod1<-glmmTMB(cbind(HATCHED, EGGS_UNBURIED-HATCHED) ~ 
                      EGGS_UNBURIED+ADULTS+as.factor(NEST_ATTMPT)+COMP+(1|SITE:LOCATION)+(1|YEAR),
                      data = ANI, 
                      family=binomial(link="logit"))

success_mod2<-glmmTMB(cbind(FLEDGED, HATCHED-FLEDGED) ~ 
                      HATCHED+ADULTS+as.factor(NEST_ATTMPT)+COMP+(1|SITE:LOCATION)+(1|YEAR),
                      data = ANI, 
                      family=binomial(link="logit"))
```

```{r}

success_mod1_b<-glmmTMB(HATCHED ~ 
                      EGGS_UNBURIED+ADULTS+as.factor(NEST_ATTMPT)+COMP+(1|SITE:LOCATION)+(1|YEAR),
                      data = ANI, 
                      family=poisson(link="log"))

success_mod2_b<-glmmTMB(FLEDGED ~ 
                      HATCHED+ADULTS+as.factor(NEST_ATTMPT)+COMP+(1|SITE:LOCATION)+(1|YEAR),
                      data = ANI, 
                      family=poisson(link="log"))
```

Summary: 

- Model 1: 

  -   There is no evidence that hatching success is impacted by clutch size
  
- Model 2: 

  -   Fledging success is negatively impacted by brood size
  

- Span Model: 

  -   There is no evidence that the clutch size impacts the hatching span
      -   Given this I did not reanalyze the first two models with the span term included as it is not significant here, and would mean using a smaller data set (<47 individuals as opposed to 121)
      -   The sub models including span would also not be able to be the "maximal" models as the others are, given that they have less data available data to estimate all the parameters I included previously. This would require dropping the "least important" parameters, and it may lead to further conflicts as what is the "least important" to me/us may not be the "least important" to others. I think/hope that this model is enough to stave off/reduce potential criticism of not including span.
  -   As the proportion of the total clutch that is incubated increases, the hatching span decreases (as competition decreases so does span). This may not be the focus of the paper, but is neat and could potentially be addressed. 

\newpage
 
#**Model 1: Hatching success as a function of clutch size**

```{r, fig.height=8}
check_model(success_mod1)

summary(success_mod1)
```

```{r}
plot(ggpredict(success_mod1, 
               terms ="EGGS_UNBURIED"))+
  geom_jitter(data=ANI, aes(EGGS_UNBURIED, (HATCHED/EGGS_UNBURIED)), 
              color="black",
              fill="black",
              stroke=0,
              width=0.15, 
              height=0.02, 
              pch=21, 
              size=3)+
  coord_cartesian(xlim = c(2.5,16))+
  scale_x_continuous( breaks=c(2,4,6,8,10,12,14,16))+
  labs(x="Clutch size", y="Hatching success", title="")+
  theme(
    axis.text = element_text(size=12, color = "black"),
    axis.ticks = element_line(color="black"),
    axis.line.x = element_line(color="black"),
    axis.line.y = element_line(color="black"),
    axis.title=element_text(color="black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    )

```

\newpage

#**Model 2: Fledging success as a fucnction of brood size**

```{r, fig.height=8}
check_model(success_mod2)

summary(success_mod2)
```

```{r}
plot(ggpredict(success_mod2, 
               terms ="HATCHED"))+
  geom_jitter(data=ANI, aes(HATCHED, (FLEDGED/HATCHED)), 
              color="black",
              fill="black",
              stroke=0,
              width=0.15, 
              height=0.02, 
              pch=21, 
              size=3)+
  coord_cartesian(xlim = c(1,12))+
  scale_x_continuous(breaks = c(2,4,6,8,10,12))+
  labs(x="Brood size", y="Fledging success", title="")+
  theme(
    axis.text = element_text(size=12, color = "black"),
    axis.ticks = element_line(color="black"),
    axis.line.x = element_line(color="black"),
    axis.line.y = element_line(color="black"),
    axis.title=element_text(color="black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    )
```

\newpage

```{r, include=F}
#PCA
#bringing in the "best first date" data and morphometrics
PCA_DAT<-read.csv("E:/quinl/Desktop/Grad School/ANI/PCA_DATA.csv")

#PCA results, and inserting PC1 back to the main dataset
PCA_DAT$PC1<-prcomp(PCA_DAT[11:13], scale = TRUE)[["x"]][,1]
theme_set(theme_few())

#PCA values for all PCs
pca_dat <- prcomp(PCA_DAT[11:13], scale = TRUE)

#looking at the R^2 of lienar models
summary(lm(pca_dat[["x"]][,1]~as.factor(PCA_DAT$DATE)))
summary(lm(PCA_DAT$DATE~PCA_DAT$TARSUS))
summary(lm(PCA_DAT$CULMEN~PCA_DAT$DATE))
summary(lm(PCA_DAT$MASS~PCA_DAT$DATE))


#Setup for Jackknife Leave one out Cross validation
library(sampling)

N=length(PCA_DAT$TARSUS)

jack.tars<-numeric(N)
jack.int1<-numeric(N)
jack.rsq1<-numeric(N)
predictions.tars<-NULL
jack.pc1<-numeric(N)
jack.int2<-numeric(N)
jack.rsq2<-numeric(N)
predictions.pc1<-NULL
jack.culm<-numeric(N)
jack.int3<-numeric(N)
jack.rsq3<-numeric(N)
predictions.culm<-NULL

for (i in 1:N) {
  
  validation<-PCA_DAT[i,]
  training<-PCA_DAT[-i,]
  
  model<-lm(DATE~TARSUS, data=training)
  jack.tars[i]<-coef(model)[2] 
  jack.int1[i]<-coef(model)[1]
  jack.rsq1[i]<-summary(model)$adj.r.squared
  predictions.tars[i]<-predict(model, newdata=validation)
  
  model<-lm(DATE~PC1, data=training)
  jack.pc1[i]<-coef(model)[2] 
  jack.int2[i]<-coef(model)[1]
  jack.rsq2[i]<-summary(model)$adj.r.squared
  predictions.pc1[i]<-predict(model, newdata=validation)
  
  model<-lm(DATE~CULMEN, data=training)
  jack.culm[i]<-coef(model)[2] 
  jack.int3[i]<-coef(model)[1]
  jack.rsq3[i]<-summary(model)$adj.r.squared
  predictions.culm[i]<-predict(model, newdata=validation)
  
}

#plotting the error associated with Leave one out cross-validation (LOOCV)
pc1.error<-(predictions.pc1-PCA_DAT$DATE)

#Generating Skew and Span Dataset from PCA
#bias corrected measures
bias_pc1<-(N*coef(lm(PCA_DAT$DATE~PCA_DAT$PC1))[2]) - (N - 1)*mean(jack.pc1)
bias_int2<-(N*coef(lm(PCA_DAT$DATE~PCA_DAT$PC1))[1]) - (N - 1)*mean(jack.int2)

#bringing in the unknown aged chicks
UNK_DATE<-read.csv("E:/quinl/Desktop/Grad School/ANI/METRICS_UNK_AGE.csv")
names(UNK_DATE) <- toupper(names(UNK_DATE))
UNK_DATE$TARSUS <- as.double(UNK_DATE$TARSUS)
UNK_DATE$MASS <- as.double(UNK_DATE$MASS)
UNK_DATE<-na.omit(UNK_DATE)
UNK_DATE$AGE<-NA

#center for scaling for PC1
UNK_DATE$TARSUS_s<-(UNK_DATE$TARSUS-mean(PCA_DAT$TARSUS))/sd(PCA_DAT$TARSUS)
UNK_DATE$CULMEN_s<-(UNK_DATE$CULMEN-mean(PCA_DAT$CULMEN))/sd(PCA_DAT$CULMEN)
UNK_DATE$MASS_s<-(UNK_DATE$MASS-mean(PCA_DAT$MASS))/sd(PCA_DAT$MASS)

#multiplying by the calculated rotations from the PCA of known data to get the PC1 value
UNK_DATE$PC1<-
  (UNK_DATE$TARSUS_s*pca_dat[["rotation"]][1])+
  (UNK_DATE$CULMEN_s*pca_dat[["rotation"]][2])+
  (UNK_DATE$MASS_s*pca_dat[["rotation"]][3])


#function to generate ages
#x is the data you want to extrapolate dates
#z is the variable you want regressed against date to generate the linear model
INTER_AGE_j<-function(coeff, data, intercept){
  w = coeff*data + intercept
  return(w)
}

#use the bias corrected estimates from the JACKKNIFE dataset
#UNK_DATE$AGE_TCONT_j <- round(INTER_AGE_j(bias_tars, UNK_DATE$TARSUS, bias_int1))                   
UNK_DATE$AGE_PCONT_j <- round(INTER_AGE_j(bias_pc1, UNK_DATE$PC1, bias_int2))
#UNK_DATE$AGE_CCONT_j <- round(INTER_AGE_j(bias_culm, UNK_DATE$CULMEN, bias_int3))
UNK_DATE<-UNK_DATE[-c(7:11)]

#trying to get data from extrapolated stuff
#reordering the dataframe so that within 
UNK_DATE<-UNK_DATE[with(UNK_DATE, order(TERRITORY,DATE, CHICK)), ]
#removing the YEAR variable to assign "Year territory" name and get hatching date
UNK_DATE$DATE <- as.Date(UNK_DATE$DATE, format = "%Y-%m-%d")
#subtracting days age from hatching date
UNK_DATE$HATCH <- UNK_DATE$DATE-(UNK_DATE$AGE_PCONT_j-1)

#giving "year territory" name
UNK_DATE$YEAR <- format(UNK_DATE$HATCH, format = "%Y")
UNK_DATE$GRP <- paste(UNK_DATE$YEAR, UNK_DATE$TERRITORY)

library(data.table)

succ<-(read.csv("E:/quinl/Desktop/Grad School/ANI/SUCCESSFUL_NESTS.csv")	%>% 
         mutate(NULL, 
                ADULTS=as.numeric(ADULTS), 
                EGGS_BURIED=as.numeric(EGGS_BURIED),
                EGGS_BURIED=as.numeric(EGGS_BURIED), 
                EGGS_UNBURIED=as.numeric(EGGS_UNBURIED),
                HATCHED=as.numeric(HATCHED), 
                FLEDGED=as.numeric(FLEDGED),
                TOT_EGGS=as.numeric(TOT_EGGS),
                LOCATION=as.factor(LOCATION),
                YEAR=as.factor(YEAR), 
                SITE=as.factor(SITE)	
         ))

#new dataset just of entries with accurate chickdates
suc<-subset(succ, SPAN>0)

#female group size instead of #adults
suc$FEM<-floor(suc$ADULTS/2)

suc$GRP<-paste(suc$YEAR, suc$LOCATION)

#from DOI : 10.1098/rspb.2018.1452
#ti = time since beginning of span
#tm = time at the middle of the span
#p = proportion at that time

skew2<-function(ti, tm, p){
  skew<-((ti-tm)/(tm))*p
  print(skew)
}

suc$SPAN<-as.numeric(suc$SPAN)
suc$skew1<-skew2(1, (suc$SPAN/2), (suc$NEW.CHCKS_DAY1/suc$HATCHED))  
suc$skew2<-skew2(2, (suc$SPAN/2), (suc$NEW.CHCKS_DAY2/suc$HATCHED))  
suc$skew3<-skew2(3, (suc$SPAN/2), (suc$NEW.CHCKS_DAY3/suc$HATCHED))  
suc$skew4<-skew2(4, (suc$SPAN/2), (suc$NEW.CHCKS_DAY4/suc$HATCHED))  
suc$skew5<-skew2(5, (suc$SPAN/2), (suc$NEW.CHCKS_DAY5/suc$HATCHED))  
suc$skew6<-skew2(6, (suc$SPAN/2), (suc$NEW.CHCKS_DAY6/suc$HATCHED))  
suc$skew7<-skew2(7, (suc$SPAN/2), (suc$NEW.CHCKS_DAY7/suc$HATCHED))  

#sums all the values to calculate skew metric
suc$SKEW<-apply(suc[23:29], 1, sum, na.rm=T)

#return to work with Unknown chick ages
#removes duplicate measures of each chick in the group and keeps the first instance
SKEW_1<-setDT(UNK_DATE)[ , .SD[(!duplicated(CHICK))], by = GRP]

SKEW_1<-SKEW_1%>%
  group_by(GRP)%>%
  summarise(DATE_DIFF=as.numeric(HATCH-min(HATCH))+1,
            SPAN=max(DATE_DIFF))
SKEW_1<-SKEW_1%>%
  group_by(GRP, DATE_DIFF, SPAN)%>%
  summarise(n=n())

#reshaping data to be able to calculate the skew values
SKEW_1<-reshape2::dcast(SKEW_1, GRP+SPAN~DATE_DIFF)

SKEW_SUCC<-succ[1:11]

SKEW_SUCC$GRP <- paste(SKEW_SUCC$YEAR, SKEW_SUCC$LOCATION)

#merging data on group success with unknown hatching data
SKEW_1<-merge(SKEW_SUCC, SKEW_1, by="GRP")

SKEW_1$skew1<-skew2(1, (SKEW_1$SPAN/2), (SKEW_1$`1`/SKEW_1$HATCHED))
SKEW_1$skew2<-skew2(2, (SKEW_1$SPAN/2), (SKEW_1$`2`/SKEW_1$HATCHED))
SKEW_1$skew3<-skew2(3, (SKEW_1$SPAN/2), (SKEW_1$`3`/SKEW_1$HATCHED))
SKEW_1$skew4<-skew2(4, (SKEW_1$SPAN/2), (SKEW_1$`4`/SKEW_1$HATCHED))
SKEW_1$skew5<-skew2(5, (SKEW_1$SPAN/2), (SKEW_1$`5`/SKEW_1$HATCHED))
SKEW_1$skew6<-skew2(6, (SKEW_1$SPAN/2), (SKEW_1$`6`/SKEW_1$HATCHED))
SKEW_1$skew7<-skew2(7, (SKEW_1$SPAN/2), (SKEW_1$`7`/SKEW_1$HATCHED))
SKEW_1$skew8<-skew2(8, (SKEW_1$SPAN/2), (SKEW_1$`8`/SKEW_1$HATCHED))
SKEW_1$skew10<-skew2(10, (SKEW_1$SPAN/2), (SKEW_1$`10`/SKEW_1$HATCHED))

SKEW_1$SKEW<-apply(SKEW_1[23:31], 1, sum, na.rm=T)

#creating empty columns so that I can merge the known and unknown datasets together without issue
suc$NEW.CHCKS_DAY8<-NA
suc$NEW.CHCKS_DAY10<-NA

#merging known and unknown aged chick datasets together
SKEW_DATA<-rbind(SKEW_1[,c("GRP", "YEAR", "LOCATION", "SITE", "ADULTS", "HATCHED", "FLEDGED", 
                           "1", "2", "3", "4",  "5", "6", "7", "8", "10", "SPAN", "SKEW")], 
                 setnames((suc[,c("GRP", "YEAR", "LOCATION", "SITE", "ADULTS", "HATCHED", "FLEDGED", "NEW.CHCKS_DAY1", 
                                  "NEW.CHCKS_DAY2", "NEW.CHCKS_DAY3", "NEW.CHCKS_DAY4", "NEW.CHCKS_DAY5", 
                                  "NEW.CHCKS_DAY6", "NEW.CHCKS_DAY7", "NEW.CHCKS_DAY8", "NEW.CHCKS_DAY10", "SPAN", "SKEW")]), 
                          names((SKEW_1[,c("GRP", "YEAR", "LOCATION", "SITE", "ADULTS", "HATCHED", "FLEDGED", 
                                          "1", "2", "3", "4",  "5", "6", "7", "8", "10", "SPAN", "SKEW")])))
)

SKEW_DATA<-SKEW_DATA%>%distinct(GRP, .keep_all=T)
SKEW_DATA$FpHa<-SKEW_DATA$FLEDGED/SKEW_DATA$HATCHED
SKEW_DATA$FEM<-floor(SKEW_DATA$ADULTS/2)

ANI$GRP<-paste(ANI$YEAR, ANI$LOCATION)
SKEW_1<-merge(SKEW_DATA, ANI, by="GRP")

SKEW_1<-SKEW_1 %>%
  mutate(Total = select(., `1`:`10`) %>% rowSums(na.rm = TRUE))
SKEW_1$TF<-SKEW_1$Total==SKEW_1$HATCHED.y
SKEW_1<-subset(SKEW_1, TF=="TRUE")
SKEW_1$COMP<-SKEW_1$EGGS_UNBURIED/SKEW_1$TOT_EGGS
```

#**Span model: Span as a function of clutch size**

```{r, include=F}
Span_mod<-glmmTMB(SPAN ~ EGGS_UNBURIED+ADULTS.x+COMP+as.factor(NEST_ATTMPT),
                  data = SKEW_1, 
                  family = poisson(link="log")
                  )
```

```{r, fig.height=8}
check_model(Span_mod)

summary(Span_mod)
```

```{r }
plot(ggpredict(Span_mod, 
               terms ="EGGS_UNBURIED"))+
  geom_jitter(data=SKEW_1, aes(EGGS_UNBURIED, SPAN), 
              color="black",
              fill="black",
              stroke=0,
              width=0.15, 
              height=0.15, 
              pch=21, 
              size=3)+
  coord_cartesian(xlim = c(4,16))+
  labs(x="Clutch size", y="Hatching Span", title="Span (days) ~                      
Clutch Size+Adults+Nest Attempt+'Competition'+Random Effects")+
  theme(
    axis.text = element_text(size=12, color = "black"),
    axis.ticks = element_line(color="black"),
    axis.line.x = element_line(color="black"),
    axis.line.y = element_line(color="black"),
    axis.title=element_text(color="black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )

```

```{r}
plot(ggpredict(Span_mod, 
                 terms ="COMP"))+
    geom_jitter(data=SKEW_1, aes(COMP, SPAN), 
                color="black",
                fill="black",
                stroke=0,
                width=.01, 
                height=0.15, 
                pch=21, 
                size=3)+
    #coord_cartesian(xlim = c(4,16))+
    labs(x="Proportion of laid clutch that is incubated", y="Hatching Span", title="Span (days) ~                      
Clutch Size+Adults+Nest Attempt+'Competition'+Random Effects")+
    theme(
      axis.text = element_text(size=12, color = "black"),
      axis.ticks = element_line(color="black"),
      axis.line.x = element_line(color="black"),
      axis.line.y = element_line(color="black"),
      axis.title=element_text(color="black"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
    )
```

